---
title: "how_to_use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how_to_use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(pttdatahaku)
```

## Browsing databases

path to database: `r pttdatahaku:::db_path`

To list all data sets in the data base, use `ptt_search_data` without arguments. To search, you can set a search term.

```{r}
ptt_search_data()
ptt_search_data("tyonv")
```

To list all data bases, use `ptt_search_database` without arguments

```{r}
ptt_search_db()
```

## Saving Data

Preferably do not save data to database manually/interactively but use the database lists for reproducibility and updating. See Section Creating Databases. Sometimes you need to save manually however, for this, use `ptt_save_data`:

```{r}
test_df <- data.frame(x = letters[1:5], y = rnorm(5))
ptt_save_data(test_df)
```

## Reading data

Use `ptt_read_data` with the name of data set to read data from a data base. 

```{r}
data <- ptt_read_data("test_df")
data
```

## Importing Data from Statistics Finland

To import data, you need an url that tells you where the data is. Function `statfi_parse_url` transforms the addresses in the GUI of pxweb to the addresses of API of pxweb:

```{r}
gui_url <- "https://pxnet2.stat.fi/PXWeb/pxweb/fi/StatFin/StatFin__kou__vkour/statfin_vkour_pxt_12bq.px/"
api_url <- statfi_parse_url(gui_url)
api_url
```

Given url, you can construct queries using `pxweb_print_full_query`:

```{r}
my_url <- api_url
pxweb_print_full_query(my_url)
```

Given url, you can print a code that imports the data in the url:

```{r, warning = FALSE}
pxweb_print_code_full_query(my_url)
pxweb_print_code_full_query_c(my_url)
```

where the latter copies the code to clipboard for you to add where ever you wish simply by ctrl+V. With a query and an url, you can import data. Let's simplify the query a bit:

```{r}
my_query <-
     list("Vuosi"=c("1970","1975"),
          "Alue"=c("SSS","KU020","KU005"),
          "Ik\U00E4"=c("SSS","15-19"),
          "Sukupuoli"=c("SSS","1","2"),
          "Koulutusaste"=c("SSS","3T8"),
         "Tiedot"=c("vaesto"))
```

Use function `ptt_get_statfi` to import data:

```{r}
ptt_get_statfi(my_url, my_query)
```

## Creating Databases

Each database has a database list. Database lists contain the code that accesses the Statistics Finland data API and imports the data to the database. For each data set there is an entry in a database list. Database lists are used to create and update databases. 

Use function `ptt_save_db_list` to create a database list. Note that we are adding an object to the data base. Thus, first create the object and then use this object as an argument of `ptt_save_db_list` without quotation marks. Database lists are list-objects.

```{r}
test_db <- list()
ptt_save_db_list(test_db)
```

To retrieve a database list, use function `ptt_read_db_list`. Note that we are fetching an object from the database by its name. Hence, use quotation marks.

```{r}
ptt_read_db_list("test_db")
```

This database is still empty. Database lists store queries. A query is a set of information that is required to import a data set. To add a query, you need the url of the data set, the query itself and a call. Query and url are the objects that are also the arguments to the `ptt_get_statfi`-function. The third element, the call, is the expression to use `ptt_get_statfi`-function itself. The call may also contain further operations as `ptt_get_statfi` cannot possibly handle all data in the wild.

To add a query to a database list, use function `ptt_db_add_query` with arguments that indicate the name of the database to which you wish to add the query, the url, and the call.

```{r}
ptt_add_query("test_db", url = my_url, query = my_query, call = c("ptt_get_statfi(url, query)"))
```

Now our test database is not empty anymore:

```{r}
db_list <- ptt_read_db_list("test_db")
db_list
```

Use R base `names`-function to see all the data sets that the database list manages:

```{r}
names(db_list)
```


Now the the database list has all information it needs to import the data. To add the data to the database, use `ptt_db_update`-function. It takes as argument the name of the database list and imports and updates all data sets on which the database list has information.

```{r}
ptt_db_update("test_db")
```

To keep better track of your data you might want to add information about your data to your database. To add metadata information from SF:

```{r}
ptt_add_pxweb_metadata("test_db")
ptt_read_db_list("test_db")$vkour_12bq$pxweb_metadata

```

You can also create complete databases using simple the url locations of required data: 

```{r, eval = FALSE}

my_new_db_list <- c("StatFin__tym__atp__nj/statfin_atp_pxt_11my.px/",
                    "https://pxnet2.stat.fi/PXWeb/api/v1/fi/StatFin/kou/vkour/statfin_vkour_pxt_12bq.px")
ptt_create_db_list(my_new_db_list, overwrite = TRUE)
names(ptt_read_db_list("my_new_db_list"))
```

To summarize a data base list:

```{r}
ptt_glimpse_db("my_new_db_list")

```

## Codes and names

```{r}
df <- ptt_read_data("tyonv_12u4")
df
```

```{r}
label_key <- get_codes_names("tyonv_12u4", "tyovoimapalvelut_db")
df <- statficlassifications::key_recode(df, label_key)
df
```

Switching between codes and names is straightforward

```{r}
df <- statficlassifications::key_recode(df, label_key)
df
```

