---
title: "how_to_use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how_to_use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pttdatahaku)
```

## Creating database lists

Database lists contain the code that accesses the SF API and imports the data to the database. Database lists can be used to create and update databases.

Use function `ptt_save_db_list` to create a database list with a name `my_database` 

```{r}
ptt_save_db_list(my_database)

```
db_lists live in the same folder as the data sets.
To retrieve a database list, use function `ptt_read_db_list`.

Database lists consist of queries. A query is a set of information that is required to import a dataset. To add a query, you need the url of the data set and the query itself. 

How to find the url?

How to construct the query? 

To add a query to a database list, use function `ptt_db_add_query`

```{r}

ptt_db_add_query("my_database", url = my_url, query = my_query)

```


## Creating new database

As a default, databases are created into a teams folder. 

Use function `ptt_db_update` to import and update data. 


## Saving data into a database

## Reading data from a database

